---
title: "Cari Assignment on Building Functions"
author: "Emily Adamic"
date: "2022-12-01"
output: 
  rmarkdown::html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

Using the ebird dataset from previous lessons, write a function that will take the scientific name as input and create a new file containing all the observation info for that bird. Name the output files with the scientific names for the following species:

```{r, include=FALSE}
library(tidyverse)
```

```{r}
df = read.csv("Data/MBT_ebird.csv")
```

```{r}

q1_func <- function(sci_name, df){
  df_output = filter(df, scientific_name == sci_name)
  csv_output = paste0("Results/",sci_name, "_output.csv")
  write.csv(df_output, csv_output)
  cat("\nFile:", csv_output, "created")
}

q1_func("Anser caerulescens", df)
q1_func("Antrostomus carolinensis", df)
q1_func("Setophaga americana", df)

```


[ANSER CAERULESCENS LINK HERE](Results/Anser caerulescens_output.csv)

[ANTROSTOMUS CAROLINENSIS LINK HERE](Results/Antrostomus carolinensis_output.csv)

[SETOPHAGA AMERICANA LINK HERE](Results/Setophaga americana_output.csv)

# Question 2

Using the ebird dataset from previous lessons, write a function that will take the scientific name as input and create a new file containing all the observation info for that bird. Name the output files with the scientific names for the following species.

**Note:** I have built a function off of the column "count", assuming these represent the individaul days that the birds were observed the most and the least? This results in a lot more rows per bird than the example (i.e. there were many days when that bird was observed only once)... but I may not understand the question perfectly, however. 

```{r}

# "Anser caerulescens"
# Antrostomus carolinensis
# Setophaga americana

q2_func <- function(sci_name) {
  
  csv_name = paste0("Results/", sci_name, "_output.csv")
  
  df = read.csv(csv_name)
  min_num = min(df$count)
  max_num = max(df$count)
  
  output_df = filter(df, count %in% c(min_num, max_num))

}

all_three_df = bind_rows(q2_func("Anser caerulescens"), q2_func("Antrostomus carolinensis"), q2_func("Setophaga americana"))

write.csv(all_three_df, "Results/q2_function_maxmin_output.csv")

```

[FILE LINK HERE](Results/q2_function_maxmin_output.csv)

# Question 3 

Write a nested function that will complete Problems 1 and 2 in a single function, but with the following species names as argument values.

```{r}
df = read.csv("Data/MBT_ebird.csv")

sci_names = c("Branta canadensis","Spatula discors", "Anas platyrhynchos")

q3_func = function(sci_names){
  
  # run q1 function on all three 
  for(s in sci_names){
    cat("\n", s)
    q1_func(s, df)
  }

  # run q2 func on all three to get the min max for each. 
  all_three_df = bind_rows(q2_func(sci_names[1]), 
                           q2_func(sci_names[2]), 
                           q2_func(sci_names[3]))
  
  write.csv(all_three_df, "Results/q3_allthree_output.csv")
  
  cat("\n\nSaved min and max observations for all in sci_names.")
  
}

q3_func(sci_names)

```


[LINK TO BRANTA CANADENSIS](Results/Branta canadensis_output.csv)

[LINK TO SPATULA DISCORS](Results/Spatula discors_output.csv)

[LINK TO ANAS PLATYRHYNCHOS](Results/Anas platyrhynchos_output.csv)

[LINK TO ALL THREE MIN AND MAX FILE](Results/q3_allthree_output.csv)




