---
title: "Madison HeatMap Assignment"
author: "Emily Adamic"
date: "2022-12-01"
output: 
  rmarkdown::html_document:
    code_folding: show
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Description: 

Take the following dataset and create a heatmap using ggplot2 that is faceted by tissue type. Customize the heatmap and make sure all the axis labels/titles are correct and that you can read them all. Make sure to add a plot title. Feel free to customize the color and whatever else you see fit. Create an R Markdown file with a link to your script as well as an imbedded image of your heatmap.

> Note: I have chosen to just do all the assignment in R Markdown, without linking to an external script. I have included the code chunk in the collapsible menu. I wasn't sure the units for expression, so I just put au. 

```{r load libs and data, include=FALSE}
library(tidyverse)
df = read.csv("Data/MAH_assignment_data.csv")
```

```{r make heatmap, fig.width=10, fig.height=6}
df %>% ggplot(aes(Sample, Gene, fill=Expression)) + 
  geom_tile(color="black", size=0.7) + 
  facet_grid(~Tissue, scales="free_x", space="free_x") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle=270, hjust = 0, family="mono"),
        axis.text.y = element_text(family="mono"),
        text = element_text(face="bold"), 
        strip.placement = "outside", 
        plot.title = element_text(hjust=0.5)) + 
  scale_x_discrete(labels=c('A.tri', 'D.brim', 'E.tyn M', 'E.tyn P', 'P.alb', 'R.kez', 
                            'A.tri', 'D.brim', 'E.tyn M', 'E.tyn P', 'P.alb', 'R.kez',
                            'A.tri', 'D.brim', 'E.tyn M', 'E.tyn P', 'P.alb', 'R.kez',
                            'A.tri', 'D.brim', 'E.tyn M', 'E.tyn P', 'P.alb', 'R.kez')) + 
  scale_fill_gradient(low = "black", high = "lime green", name="Expression (au)") + 
  labs(x="", y="Gene", title = "Gene Expression Across Tissues")

```





